---
title: "Human Activity Recognition"
author: "Deepan Sarkar"
date: "Sunday, November 23, 2014"
output: html_document
---

## Load the data set and pre-process the data
  
### Loading the data set
```{r, echo=TRUE, cache=TRUE}
library( caret )
library( randomForest )
train <- read.csv( "E:/COURSES/DATA SCIENCE/8 Practical Machine Learning/Assignment/pml-training.csv" )
test <- read.csv( "E:/COURSES/DATA SCIENCE/8 Practical Machine Learning/Assignment/pml-testing.csv" )
```
  
### Pre-processing the data set
Here I have considered only 52 variables as the predictor variables. All these variables have no missing data values.  
The train data has been further divided into training data set and testing data set by random 75%-25% split. This is done to validate whether the generated model is accurate enough to be used on the test data set.
```{r, echo=TRUE, cache=TRUE}
Y <- "classe"
Xs <- c( "roll_belt", "pitch_belt", "yaw_belt", "total_accel_belt", "gyros_belt_x", "gyros_belt_y", "gyros_belt_z", "accel_belt_x", "accel_belt_y", "accel_belt_z", "magnet_belt_x", "magnet_belt_y", "magnet_belt_z", "roll_arm", "pitch_arm", "yaw_arm", "total_accel_arm", "gyros_arm_x", "gyros_arm_y", "gyros_arm_z", "accel_arm_x", "accel_arm_y", "accel_arm_z", "magnet_arm_x", "magnet_arm_y", "magnet_arm_z", "roll_dumbbell", "pitch_dumbbell", "yaw_dumbbell", "total_accel_dumbbell", "gyros_dumbbell_x", "gyros_dumbbell_y", "gyros_dumbbell_z", "accel_dumbbell_x", "accel_dumbbell_y", "accel_dumbbell_z", "magnet_dumbbell_x", "magnet_dumbbell_y", "magnet_dumbbell_z", "roll_forearm", "pitch_forearm", "yaw_forearm", "total_accel_forearm", "gyros_forearm_x", "gyros_forearm_y", "gyros_forearm_z", "accel_forearm_x", "accel_forearm_y", "accel_forearm_z", "magnet_forearm_x", "magnet_forearm_y", "magnet_forearm_z" )
set.seed(12345)
inTrain <- createDataPartition( y=train$classe, p=0.75, list=FALSE )
training <- train[inTrain,c(Xs,Y)]
testing <- train[-inTrain,c(Xs,Y)]
```

## Generate model from the data
  
### The model
The training data set is used to build a random forest model.
```{r, echo=TRUE, cache=TRUE}
fit<-randomForest(classe~.,training)
predTrain <- predict( fit, training )
confTrain <- confusionMatrix( predTrain, training$classe )
```
The out of sample error is expected to be around `r (1-confTrain$overall[1])*100`%.
  
### The accuracy
The prediction accuracy on the testing data set is shown below.
```{r, echo=TRUE, cache=TRUE}
predTest <- predict( fit, testing )
confTest <- confusionMatrix( predTest, testing$classe )
```
The out of sample error is found to be around `r (1-confTest$overall[1])*100`%.

### The predictions
The predictions on the test data set is given below.
```{r, echo=TRUE, cache=TRUE}
predict( fit, test )
```

## Citation
The data is obtained from Human Activity Recognition data set at the url: http://groupware.les.inf.puc-rio.br/har.
